---
title: "HW 4 - AirPassengers Time Series Analysis"
output: html_document
author: "Aryan Walia"
date: "2025-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(ggplot2)
library(forecast)
library(tseries)
library(TTR)
library(zoo)

data <- read.csv("~/Downloads/AirPassengers.csv", header = TRUE, check.names = TRUE)
data_ts <- ts(data$Passengers, start = c(1949, 1), frequency = 12)

```{r setup, include=FALSE}
```{r plot_time_series, echo=FALSE}
autoplot(data_ts) + 
  ggtitle("Air Passengers Time Series") + 
  ylab("Number of Passengers") + 
  xlab("Year") +
  theme_minimal()

decomposed <- decompose(data_ts, type = "multiplicative")
autoplot(decomposed) + ggtitle("Decomposed Air Passengers Time Series")

acf(data_ts, main = "Autocorrelation of AirPassengers Data")

data_diff <- diff(log(data_ts))
adf_result <- adf.test(data_diff)
print(paste("ADF Test p-value:", adf_result$p.value))

arima_model <- auto.arima(data_ts)
summary(arima_model)

ets_model <- ets(data_ts)
summary(ets_model)

arima_forecast <- forecast(arima_model, h = 12)
ets_forecast <- forecast(ets_model, h = 12)

autoplot(data_ts) +
  autolayer(arima_forecast$mean, series = "ARIMA Forecast", PI = TRUE) +
  autolayer(ets_forecast$mean, series = "ETS Forecast", PI = TRUE) +
  ggtitle("Forecasts from ARIMA and ETS") +
  ylab("Number of Passengers") +
  xlab("Year") +
  theme_minimal()

arima_accuracy <- accuracy(arima_forecast)
ets_accuracy <- accuracy(ets_forecast)

# Print accuracy tables
print(arima_accuracy)
print(ets_accuracy)


# Extract RMSE
arima_rmse <- arima_accuracy[nrow(arima_accuracy), "RMSE"]
ets_rmse <- ets_accuracy[nrow(ets_accuracy), "RMSE"]

comparison <- data.frame(
  Model = c("ARIMA", "ETS"),
  RMSE = c(arima_rmse, ets_rmse)
)

print(comparison)


best_model <- comparison[which.min(comparison$RMSE), ]
print(paste("The best model based on RMSE is:", best_model$Model))






