---
title: "hw5"
author: "Aryan walia"
output: html_document
date: "2025-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(ggplot2)

# Import the dataset
data <- read.csv("AirPassengers.csv")

# Convert to time series
ts_data <- ts(data$Passengers, start = c(1949, 1), frequency = 12)
```

```{r}
autoplot(ts_data) + ggtitle("Monthly Air Passengers") + ylab("Passengers")
```
```{r}
model <- auto.arima(ts_data)
summary(model)
checkresiduals(model)
```
### 1. Explaining the Output

The model selected by `auto.arima()` is **ARIMA(2,1,1)(0,1,0)[12]**, which means:

- **p = 2**: Two autoregressive terms (AR).
- **d = 1**: First-order differencing to remove trend.
- **q = 1**: One moving average term (MA).
- **Seasonal (0,1,0)[12]**: One seasonal differencing applied, no seasonal AR/MA terms.

The coefficients section shows the estimated values for ar1, ar2, and ma1, with standard errors.  
The model fit is measured by:

- **AIC = 1017.85**, **BIC = 1029.35**: Lower values indicate a better model.
- **sigma² = 132.3**: The variance of the residuals; lower means better fit.

These values suggest the model fits the data well and captures both trend and seasonality.

```{r}
forecasted <- forecast(model, h = 5)
autoplot(forecasted) + ggtitle("Forecast for Next 5 Periods") + ylab("Passengers")
```
```{r}
accuracy(model)
```



