---
title: "Forecasting US Arrivals to Australia"
author: "Aryan Walia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data

```{r load-packages}
library(fpp2)

```

```{r load-data}
data("arrivals")
ts_data <- arrivals[, "US"]  # Extract US arrivals
ts_data <- ts(ts_data, start = c(1981, 1), frequency = 12)
```

## Time Series Plot

```{r plot-time-series}
ggplot2::autoplot(ts_data) +
  ggtitle("US Arrivals to Australia") +
  xlab("Year") +
  ylab("Number of Arrivals")
```

**Observations:**
- There appears to be an increasing trend over time.
- There is evident seasonality, with peaks and dips occurring at regular intervals.
- Some fluctuations are more extreme in certain periods.

## Summary Statistics

```{r summary-stats}
summary(ts_data)
boxplot(ts_data, main="Boxplot of US Arrivals to Australia", ylab="Arrivals")
```

**Observations:**
- The mean and median are relatively close, indicating a somewhat symmetric distribution.
- There are potential outliers in the upper range.
- Seasonality likely contributes to the variations in values.

## Time Series Decomposition

```{r decomposition}
decomp <- decompose(ts_data, type = "multiplicative")
plot(decomp)
```

**Observations:**
- The series exhibits a clear seasonal pattern.
- The decomposition appears to be **multiplicative** since variations increase over time.
- Seasonality indices reveal months with high and low arrivals.

```{r seasonal-indices}
decomp$seasonal[1:12]  # Extract seasonal indices
```

- The highest arrivals occur in **December** (holiday season).
- The lowest arrivals occur in **February**, likely due to fewer vacations.

## Seasonality-Adjusted Plot

```{r seasonality-adjusted}
adjusted_ts <- ts_data / decomp$seasonal
autoplot(ts_data, series="Actual") +
  autolayer(adjusted_ts, series="Seasonally Adjusted", color='red') +
  ggtitle("Seasonally Adjusted vs. Actual US Arrivals")
```

**Observations:**
- The adjusted time series smooths out seasonal fluctuations.

## Naïve Method Forecast

```{r naive-forecast}
naive_model <- naive(ts_data)
plot(naive_model)
```

## Residual Analysis

```{r residual-analysis}
residuals <- residuals(naive_model)

# Plot Residuals
autoplot(residuals) + ggtitle("Residuals Plot")

# Histogram
hist(residuals, main="Histogram of Residuals", xlab="Residuals", col="lightblue")

# Fitted vs Residuals
plot(fitted(naive_model), residuals, main="Fitted vs Residuals", xlab="Fitted", ylab="Residuals")
abline(h=0, col="red")

# Actual vs Residuals
plot(ts_data, residuals, main="Actual vs Residuals", xlab="Actual", ylab="Residuals")
abline(h=0, col="red")

# ACF Plot
Acf(residuals)
```

**Observations:**
- Residuals appear to have patterns, suggesting some autocorrelation.
- The histogram suggests normality but with slight skewness.
- The ACF plot indicates potential correlation in residuals.

## Forecasting Accuracy

```{r accuracy-metrics}
accuracy(naive_model)
```
## Final Forecast
```{r}
naive_forecast <- naive(arrivals[,4])
plot(naive_forecast)

plot(residuals(naive_forecast))

hist(residuals(naive_forecast))

plot(fitted(naive_forecast), residuals(naive_forecast))

plot(naive_forecast$x, residuals(naive_forecast))

Acf(residuals(naive_forecast), main="ACF Plot of Residuals")



### **Summary, Accuracy, and Prediction**

summary(naive_forecast)

accuracy(naive_forecast)

forecast(naive_forecast, h=4)

plot(forecast(naive_forecast, h=4))


##### The model demonstrates low accuracy. The projected time series value for the next year is approximately 106.54, with errors increasing as the forecast extends further into the future.
```

```